{Gravitational wave sources in our Galactic backyard:}\\{Predictions for BHBH, BHNS and NSNS binaries detectable with LISA

### 2022-07-06 paper/tex/sections/variations.tex
Updates from Overleaf

Nearly all systems that can be detected form through channels where the very first interaction is stable mass transfer. Generally, higher mass transfer efficiencies lead to higher masses for the accreting stars, but also leads to wider orbits \citep{Soberman+1997, vanSon+2020}. Changing $\beta$ thus already affects the masses and orbital separation after the first interaction phase, which in turn changes the starting conditions and outcome of all subsequent interactions phases. This makes it complicated to fully understand the impact of varying $\beta$ in simple terms, but we can distinguish two main patterns for higher and lower mass systems respectively.

@@

\begin{figure*}[p]
    \centering
    \includegraphics[width=\textwidth]{fig8_dco_detections.pdf}
    \caption{The number of expected detections in the LISA mission for different DCO types and model variations. Error bars show the 1- (solid) and 2-$\sigma$ (dotted) Poisson uncertainties. An arrow indicates that the error bar extends to zero. The left axis and grid lines show the number of detections in a 4-year LISA mission and the right axis shows an approximation of the number of detections in a 10-year mission (we scale the axis by $\sqrt{T_{\rm obs}}$, see Table~\ref{tab:detection_rates} for exact rates). Each model is described in further detail in Table~\ref{tab:physics_variations} and details of the fiducial assumptions are in Section~\ref{app:fiducial_physics}. See Sec.~\ref{sec:detection_rate_analysis} for a discussion. \href{https://github.com/TomWagg/detecting-DCOs-in-LISA/blob/main/paper/figures/fig8_dco_detections.pdf}{\faFileImage} \href{https://github.com/TomWagg/detecting-DCOs-in-LISA/blob/main/paper/figure_notebooks/detections.ipynb}{\faBook}.}
    \label{fig:detection_rates}
\end{figure*}

@@

\subsubsection{Using LISA to investigate the lower mass gap}\label{sec:lower_mass_gap}

@@

From the stark difference between these models, it is clear that it is difficult at this point to say with any certainty what fraction of systems LISA will detect in the lower mass gap given the highly uncertain formation rate of systems in this mass range. Indeed, we find that the percentage of detectable systems with at least one component in the lower mass gap varies between approximately 30-70\% (or, in terms of detections, from 15 to 156) for the different model variations. However, it is important to highlight that \textit{if} DCOs are formed with components in the lower mass gap, LISA \textit{will} be able to detect them. And thus, LISA could be a useful instrument for providing constraints on the existence or non-existence of a lower mass gap based on the mass distribution of detected DCOs.

>>>>

Since nearly all systems that can be detected form through channels where the very first interaction is stable mass transfer, changing $\beta$ can affect the majority of systems. Generally, higher mass transfer efficiencies lead to higher masses for the accreting stars, but also lead to wider orbits \citep{Soberman+1997, vanSon+2020}. Changing $\beta$ thus already affects the masses and orbital separation after the first interaction phase, which in turn changes the starting conditions and outcome of all subsequent interactions phases. This makes it complicated to fully understand the impact of varying $\beta$ in simple terms, but we can distinguish two main patterns for higher and lower mass systems respectively.

@@

\begin{figure*}[p]
    \centering
    \includegraphics[width=\textwidth]{fig8_dco_detections.pdf}
    \caption{The number of expected detections in the LISA mission for different DCO types and model variations. Error bars show the 1- (solid) and 2-$\sigma$ (dotted) Poisson uncertainties. An arrow indicates that the error bar extends to zero. The left axis and grid lines show the number of detections in a 4-year LISA mission and the right axis shows an approximation of the number of detections in a 10-year mission (we scale the axis by $\sqrt{T_{\rm obs}}$, see Table~\ref{tab:detection_rates} for exact rates). \edit1{The grey shaded regions indicate the 1- and 2-$\sigma$ region for the fiducial model.} Each model is described in further detail in Table~\ref{tab:physics_variations} and details of the fiducial assumptions are in Section~\ref{app:fiducial_physics}. See Sec.~\ref{sec:detection_rate_analysis} for a discussion. \href{https://github.com/TomWagg/detecting-DCOs-in-LISA/blob/main/paper/figures/fig8_dco_detections.pdf}{\faFileImage} \href{https://github.com/TomWagg/detecting-DCOs-in-LISA/blob/main/paper/figure_notebooks/detections.ipynb}{\faBook}.}
    \label{fig:detection_rates}
\end{figure*}

@@

\subsubsection{Effect of remnant mass prescription on lower mass gap}\label{sec:lower_mass_gap}

@@

From the stark difference between these models, it is clear that it is difficult at this point to say with any certainty what fraction of systems LISA will detect in the lower mass gap given the highly uncertain formation rate of systems in this mass range.

We find that the percentage of detectable systems with at least one component in the lower mass gap varies between approximately 30-70\% (or, in terms of detections, from 15 to 156) for the different model variations that we consider, except for model \modRapid{}, for which it is 0\%. \edit1{We note however that the masses of detectable systems are not always well constrained (see Section~\ref{sec:measurement_uncertainties}). It may therefore be difficult to assess whether a detected system truly has components in the mass gap.}

\edit1{Lastly, we highlight that recent investigations of the GW detections so far suggest a significant reduction of BHs in the lower mass gap \citep[e.g. ][van Son et al.\ in prep.]{Farah+2021, Ye+2022}. All our models except model \modRapid{} allow for systems in the mass gap, which may be at tension with the observational constraints so far.}
